A systematic comparison of interaction models for systems of cities

### 2018-10-02 Docs/CCS18/Communication/ccs2018.tex

\date{CCS 2018\\\smallskip
Thessaloniki\\\smallskip
September 24th 2018
}

@@

% Understanding patterns of growth for cities is a crucial issue for its practical and policy implications e.g. regarding sustainability, but also from a theoretical point of view regarding the validation of theories for urban systems. A particular entry is taken by the Evolutive Urban Theory [1] which postulates interactions between cities as main drivers of their growth. Within this framework, numerous concurrent models have been introduced for the evolution of systems of cities, and applied on diverse case studies, but there is to the best of our knowledge no systematic comparison of performances across models and application cases. This contribution makes a first step towards such a systematic benchmark. We consider simple models simulating population of cities only, but including heterogenous underlying processes and assumptions. More precisely, we compare (i) the Favaro-Pumain model for the diffusion of innovation [2]; (ii) the Marius multi-model based on economic exchanges [3]; and (iii) a model including flows within abstract physical networks parametrized on elevation data [4]. These models are calibrated on a large scale harmonized dataset [5] including the European, former Soviet Union, Chinese, Brazilian, South-African, Indian and USA systems of cities on a time period covering 1960-2010. Calibrations are done for different versions of each model including more or less parameters, using distributed genetic algorithms on a computation grid through the intermediary of the OpenMOLE model exploration software [6]. Calibration results show that: (i) at a fixed number of parameters, no model performs particularly better on all urban systems, suggesting the complementary of the innovation, economic, and network dimensions taken into account by the different models; and (ii) when adjusting for the number of parameters with an empirical information criterion [4], we find that additional components generally improve performances for all models, what reveals a high effective dimensionality of urban systems. This work confirms the need for multiple complementary approaches to model urban systems and sketches a framework allowing systematic benchmarks of concurrent models for complex urban systems.

@@

@@

@@

@@

@@

@@

@@

Harmonized dataset, explored in particular by \cite{pumain2015multilevel} (built in the context of Geodivercity ERC)

@@

@@

@@

\sframe{Results: empirical AIC}{
% TODO : select optimal points / run empirical AIC
% OR : run on all points of the Pareto front ? -> local OK

@@

@@

>>>>

\date{CCS 2018\\\smallskip
\textit{Session Infrastructure, Planning and Environment}\\\smallskip
Thessaloniki\\\smallskip
September 24th 2018
}

@@

@@

% Working in geography has this advantage that we always situate in an interesting object of study, and reflexivity is therin inevitable.
% I have taken the habit to open my talks with some local context, and this one will not fail this rule, especially because of the beauty and hostory of Thessaloniki.
% Urban systems are fascinating as they are truly multi-scalar, both in time and space. In particular, they have a very long temporcal memory, partly because of the pphysical artifats : here the Rotunda Galerius which was later a chruch and a Mosque, or the white tower constructed by the ottoman empire ; but also because of non-ergodic dynamics.
% In this geo-hostorical context, understanding urban growth is a real challenge : to what extent the different relation of a given city influence its current growth and its future trajectory ?

@@

% We will switch to shorter time scales, but still remain at macroscopic and mesoscopic scales.

% Understanding patterns of growth for cities is a crucial issue for its practical and policy implications e.g. regarding sustainability, but also from a theoretical point of view regarding the validation of theories for urban systems.

@@

% A particular entry is taken by the Evolutive Urban Theory [1] which postulates interactions between cities as main drivers of their growth.

@@

% Within this framework, numerous concurrent models have been introduced for the evolution of systems of cities, and applied on diverse case studies, but there is to the best of our knowledge no systematic comparison of performances across models and application cases. This contribution makes a first step towards such a systematic benchmark.

@@

%  We consider simple models simulating population of cities only, but including heterogenous underlying processes and assumptions. More precisely, we compare (i) the Favaro-Pumain model for the diffusion of innovation [2]; (ii) the Marius multi-model based on economic exchanges [3]; and (iii) a model including flows within abstract physical networks parametrized on elevation data [4].

@@

% These models are calibrated on a large scale harmonized dataset [5] including the European, former Soviet Union, Chinese, Brazilian, South-African, Indian and USA systems of cities on a time period covering 1960-2010.

@@

Harmonized dataset, explored in particular by \cite{pumain2015multilevel} (built in the context of Geodivercity ERC): Urban systems for Europe, United States, Brazil, China, India, South Africa, Russia, consistent and comparable between 1960 and 2010 each 10 years.

@@

% Calibrations are done for different versions of each model including more or less parameters, using distributed genetic algorithms on a computation grid through the intermediary of the OpenMOLE model exploration software [6].

@@

%Calibration results show that: (i) at a fixed number of parameters, no model performs particularly better on all urban systems, suggesting the complementary of the innovation, economic, and network dimensions taken into account by the different models.

@@

\sframe{Results: empirical AIC}{

% (ii) when adjusting for the number of parameters with an empirical information criterion [4], we find that additional components generally improve performances for all models, what reveals a high effective dimensionality of urban systems.

% - select optimal points / run empirical AIC
% OR : run on all points of the Pareto front ? -> local OK

@@

% This work confirms the need for multiple complementary approaches to model urban systems and sketches a framework allowing systematic benchmarks of concurrent models for complex urban systems.

@@

\sframe{Models settings}{

$\rightarrow$ Work under Gibrat independence assumptions, i.e. $\Covb{P_i(t)}{P_j(t)}=0$. If $\vec{P}(t+1)=\mathbf{R}\cdot \vec{P}(t)$ where $\mathbf{R}$ is also independent, then $\Eb{\vec{P}(t+1)}=\Eb{\mathbf{R}}\cdot\Eb{\vec{P}}(t)$. Consider expectancies only (higher moments computable similarly)

$\rightarrow$ With $\vec{\mu}(t)=\Eb{\vec{P}(t)}$, we generalize this approach by taking $\vec{\mu}(t+1)=f(\vec{\mu}(t))$

\sframe{Network model}{

Direct network interaction model \cite{raimbault2018indirect}:

Let $\vec{\mu}(t)=\Eb{\vec{P}(t)}$ cities population and $(d_{ij})$ distance matrix

Model specified by

\[
f(\vec{\mu}) = r_0\cdot \mathbf{Id}\cdot \vec{\mu} + \mathbf{G}\cdot \mathbf{1} + \mathbf{N}
\]

 with 
\begin{itemize}
\item $G_{ij} = w_G\cdot \frac{V_{ij}}{<V_{ij}>}$ and $V_{ij} = \left(\frac{\mu_i\mu_j}{\sum{\mu_k}^2}\right)^{\gamma_G} \exp{(-d_{ij}/d_G)}$
\item $N_{i} = w_N \cdot \sum_{kl} \left(\frac{\mu_k\mu_l}{\sum\mu}\right)^{\gamma_N}\exp{(-d_{kl,i})/d_N}$ where $d_{kl,i}$ is distance to shortest path between $k,l$ computed with slope impedance ($Z=\left(1+\alpha/\alpha_0\right)^{n_0}$ with $\alpha_0\simeq 3$)
\end{itemize}

\sframe{Innovation diffusion}{

Favaro-Pumain model \cite{favaro2011gibrat}:

1) Diffuse innovations according to 
\[
\delta_{c,i,t} = \frac{\sum_j p_{c,j,t-1}^{s_c} \exp{(-\lambda_s d_{ij})}}{\sum_c \sum_j p_{c,j,t-1}^{s_c} \exp{(-\lambda_s d_{ij})}}
\]

2) Update population with $G_{ij}$ (see network model) such that
\[
V_{ij}= \frac{p_i p_j}{(\sum_k p_k)^2} \exp{(-\lambda_m d_{ij} \prod_c \delta_{c,i}^{\phi_c})}
\]
with $\phi_c = \sum_i p_{i,c}/\sum_{i,c} p_{i,c}$

3) Introduce innovation with utility $s_{c+1} = g_0 \cdot s_c$ in a randomly chosen city with a hierarchy parameter $\alpha_I$, if global adoption share $\phi_c$ is larger than a threshold $\theta_I$. Initial utility $s_0$ is a parameter. New innovation has an initial penetration rate $r_I$ in the city.

\sframe{Economic exchanges}{

Marius model \cite{10.1371/journal.pone.0138212}:

Initial wealth as a power law of population (exponent $\alpha_W$)

1) Update supply and demands as superlinear functions of population (exponents $\alpha_S,\alpha_D$)

2) Exchange goods according to a gravity potential of interaction (distance decay $d_M$), supplies and demands; update wealth accordingly

3) Update population such that population difference is a power law of wealth difference (economic multiplier $e_M$ and exponent $\alpha_P$)

\sframe{Benchmarked models}{

\begin{enumerate}
	\item Gibrat model: 1 param. $r_0$
	\item Direct interaction model (geographical distance): 4 param. $r_0,w_G,\gamma_G,d_G$
	\item Physical network interaction model (topographical distance: 4 param. $r_0,w_G,\gamma_G,d_G$
	\item Innovation diffusion model (simplified): 4 param. $r_0,w_I,\lambda_s,\lambda_m$ (other parameters at default values from \cite{favaro2011gibrat})
	\item Innovation diffusion model (full): 9 param. $r_0,w_I,\lambda_s,\lambda_m,s_0,g_0,r_I,\alpha_I,\theta_I$
	\item Restricted Marius model: 4 param. $e_M,\alpha_S,\alpha_D,d_M$
	\item Marius model: 6 param. $e_M,\alpha_S,\alpha_D,d_M,\alpha_W,\alpha_P$
\end{enumerate}
